### 2024-12-03 paper/paper.qmd
some updates. still need: paper and datasheet

---
title: "How do lifestyle and poverty influence BMI across different age groups and genders"
subtitle: "Negative Correlations with Poverty and Physical Activity, Positive Correlations with Age, Sleep Duration, and Gender"
author: 
  - Sakura Hu
thanks: "Code and data are available at: [https://github.com/xycww/BMI](https://github.com/xycww/BMI)."
date: today
date-format: long
abstract: "This paper investigates the relationship between BMI and factors such as lifestyle, poverty, age, and gender, aiming to identify patterns that influence maintaining a healthy BMI. The analysis uses the NHANES dataset from the US National Health and Nutrition Examination Survey and applies a multilinear regression model. The findings suggest weak overall correlations between log(BMI) and the predictors, though some significant patterns are observed: BMI is negatively associated with poverty and physical activity levels and positively associated with age, sleep duration, and being male. These results underscore the complexity of factors influencing BMI and highlight potential areas for targeted public health interventions to promote healthier lifestyles."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
library(patchwork)
library(arrow)
library(broom)
```

# Introduction

Body mass index (BMI) is a widely used measure for assessing whether an individual's weight is within a healthy range, with significant implications for health outcomes such as heart disease, diabetes, and mortality. Given its importance, understanding the factors that influence BMI, such as socioeconomic and lifestyle variables, has become an essential area of public health research. This paper aims to address the relationship about how variables like poverty, physical activity, age, sleep duration, and gender interact to influence BMI outcomes.

Using data from the US National Health and Nutrition Examination Survey (NHANES), this study models log-transformed BMI as a function of five predictors: poverty level, physical activity frequency (measured in days), age, sleep duration, and gender. A multilinear regression approach was employed to quantify these relationships and identify patterns within the dataset.

The results indicate that while the correlations between BMI and these predictors are generally weak, several significant relationships are observed. Poverty is associated with a 0.01 decrease in log(BMI), suggesting that higher income corresponds to lower BMI levels. Physical activity frequency also demonstrates a negative relationship with BMI; each additional day of physical activity per week is associated with a 0.004 decrease in log(BMI), indicating a modest benefit of regular exercise for maintaining lower BMI. Sleep duration shows a negative relationship, where one additional hour of sleep per night is linked to a 0.01 decrease in log(BMI). In contrast, age is positively associated with BMI, with each additional year corresponding to a 0.001 increase in log(BMI), reflecting the gradual weight gain commonly seen with aging. Lastly, being male is associated with a 0.02 increase in log(BMI) compared to females, suggesting possible physiological or behavioral differences between genders.

These results are essential for informing public health initiatives aimed at addressing weight-related health challenges. By identifying specific socioeconomic and lifestyle factors that influence BMI, this research provides a foundation for developing targeted interventions and strategies to promote healthier weight maintenance.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

The dataset used in this analysis is derived from the US National Health and Nutrition Examination Survey (NHANES), version 2.1.0, published in July 2015. This version of the dataset includes survey results collected between October 2009 and December 2010. NHANES is a long-running study conducted by the US National Center for Health Statistics (NCHS) that has been gathering health and nutrition data since the early 1960s. Since 1999, approximately 5,000 individuals from various age groups have been interviewed annually in their homes and undergone health examinations at mobile examination centers (MEC). 
Two datasets in this package were considered for this analysis: NHANESraw, which is the original raw data, and NHANES, a resampled version of the NHANES data. NHANESraw contains the original survey data with 20,293 observations and additional variables describing the sample weighting scheme, while NHANES is a simplified version with 10,000 resampled observations to account for oversampling effects. NHANES is used in this analysis due to its ability to reduce the potential biases from the complex survey design in NHANESraw. Additional details about this choice are provided in Appendix A. The data used here was originally compiled by Michelle Dalrymple from Cashmere High School and Chris Wild from the University of Auckland for educational purposes.

For the current study, the data was cleaned to focus on variables pertinent to the analysis of BMI. Specifically, variables such as BMI, poverty index, physical activity days, sleep hours, gender, and age were retained. After cleaning the missing values in the dataset, 3,573 observations remained.

The dataset was prepared, cleaned, and analyzed using R [@citeR] with the following libraries: tidyverse [@tidyverse] and dplyr [@dplyr] for data manipulation, ggplot2 [@ggplot2] and patchwork [@patchwork] for visualizations, and broom [@broom] for model summaries. The arrow [@arrow] library was used for efficient data storage and retrieval, while knitr [@knitr] facilitated report generation. The NHANES package [@nhanes] provided access to the dataset, and styler [@styler] was employed to ensure well-structured R code. Additionally, testthat [@testthat] and pointblank [@pointblank] were utilized for data validation and testing.

A summary table of cleaned data is shown in table 1.

```{r}
#| echo: false
#| label: table-summary
#| tbl-cap: "Summary statistics for variables in the NHANES dataset."

# Load the data
nhanes_data <- read_parquet("/Users/hxw/marriage/data/02-analysis_data/clean_data.parquet")

# Create summary statistics for each variable
summary_table <- nhanes_data %>%
  reframe(
    Variable = c(
      "BMI",
      "log(BMI)",
      "Poverty Index",
      "Physical Activity Days",
      "Age",
      "Sleep Hours"
    ),
    Mean = round(c(
      mean(BMI),
      mean(log(BMI)),
      mean(Poverty),
      mean(PhysActiveDays),
      mean(Age),
      mean(SleepHrsNight)
    ), 3),
    Median = round(c(
      median(BMI),
      median(log(BMI)),
      median(Poverty),
      median(PhysActiveDays),
      median(Age),
      median(SleepHrsNight)
    ), 3),
    Min = round(c(
      min(BMI),
      min(log(BMI)),
      min(Poverty),
      min(PhysActiveDays),
      min(Age),
      min(SleepHrsNight)
    ), 3),
    Max = round(c(
      max(BMI),
      max(log(BMI)),
      max(Poverty),
      max(PhysActiveDays),
      max(Age),
      max(SleepHrsNight)
    ), 3),
    `1st Quantile` = round(
      c(
        quantile(BMI, 0.25),
        quantile(log(BMI), 0.25),
        quantile(Poverty, 0.25),
        quantile(PhysActiveDays, 0.25),
        quantile(Age, 0.25),
        quantile(SleepHrsNight, 0.25)
      ),
      3
    ),
    `3rd Quantile` = round(
      c(
        quantile(BMI, 0.75),
        quantile(log(BMI), 0.75),
        quantile(Poverty, 0.75),
        quantile(PhysActiveDays, 0.75),
        quantile(Age, 0.75),
        quantile(SleepHrsNight, 0.75)
      ),
      3
    )
  )

# Create the summary table using kable
summary_table %>%
  kable()
```

```{r}
#| warning: false
#| message: false
#| label: fig-bmi-logbmi
#| fig-cap: "Distributions of BMI and log(BMI) in the NHANES dataset."
#| echo: false
#| fig.width: 8
#| fig.height: 4

# Create histograms with customized text sizes
hist_bmi <- ggplot(nhanes_data, aes(x = BMI)) +
  geom_histogram(binwidth = 1.5,
                 fill = "lightblue",
                 color = "black") +
  labs(title = "BMI Distribution", x = "BMI", y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

hist_log_bmi <- ggplot(nhanes_data, aes(x = log(BMI))) +
  geom_histogram(binwidth = 0.1,
                 fill = "lightblue",
                 color = "black") +
  labs(title = "log(BMI) Distribution", x = "log(BMI)", y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

# Figure 1: BMI and log(BMI)
figure_bmi <- (hist_bmi | hist_log_bmi)
figure_bmi

```

```{r}
#| warning: false
#| message: false
#| label: fig-predictor-distributions
#| fig-cap: "Distributions of predictor variables in the NHANES dataset, including Poverty, Physical Activity Days, Age, Sleep Hours, and Gender."
#| echo: false
#| fig.width: 8
#| fig.height: 12

hist_poverty <- ggplot(nhanes_data, aes(x = Poverty)) +
  geom_histogram(
    binwidth = 0.8,
    fill = "lightgreen",
    color = "black"
  ) +
  labs(title = "Poverty Index Distribution", x = "Poverty Index", y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

hist_physactive <- ggplot(nhanes_data, aes(x = PhysActiveDays)) +
  geom_histogram(
    binwidth = 1,
    fill = "lightcoral",
    color = "black"
  ) +
  labs(title = "Physical Activity Days Distribution", x = "Physical Activity Days", y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )

hist_age <- ggplot(nhanes_data, aes(x = Age)) +
  geom_histogram(
    binwidth = 3.5,
    fill = "lightyellow",
    color = "black"
  ) +
  labs(title = "Age Distribution", x = "Age", y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )

hist_sleep <- ggplot(nhanes_data, aes(x = SleepHrsNight)) +
  geom_histogram(binwidth = 1,
                 fill = "lightpink",
                 color = "black") +
  labs(title = "Sleep Hours Distribution", x = "Sleep Hours", y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )

hist_gender <- ggplot(nhanes_data, aes(x = Gender)) +
  geom_bar(fill = "lightgray", color = "black") +
  labs(title = "Gender Distribution", x = "Gender", y = "Count") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8)
  )

# Figure 2: Predictor Variables
figure_predictors <- (hist_poverty | hist_physactive | hist_age) /
  (hist_sleep | hist_gender) +
  plot_annotation(caption = "Figure 2: Distributions of predictor variables in the NHANES dataset, including Poverty, Physical Activity Days, Age, Sleep Hours, and Gender.",
                  theme = theme(plot.caption = element_text(hjust = 0.5, size = 12)))
figure_predictors

```

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

The primary outcome variable in this study is log-transformed BMI, a measure of body mass index adjusted to normalize its distribution. BMI is widely used to assess healthy weight relative to height, and its relevance to health outcomes such as cardiovascular disease and diabetes has been well-established. For this analysis, BMI was log-transformed to address its skewed distribution shown in Figure 1, providing a better fit for statistical modeling.

## Predictor variables
The following predictor variables were examined to assess the potential lifestyle and socio-economic factors influencing BMI:

- `Poverty`: This variable represents the ratio of a family’s income to the federal poverty guidelines, with lower values indicating higher levels of poverty.
- `Physical Activity Days (PhysActiveDays)`: The number of days in a typical week that a participant engages in moderate or vigorous physical activity. This variable is recorded for individuals aged 12 years and older.
- `Sleep Duration (SleepHrsNight)`: The self-reported average number of hours of sleep a participant receives on weekdays or workdays. This variable is recorded for individuals aged 16 years and older.
- `Gender`: The gender of the participant, categorized as male or female.
- `Age`: The participant's age at the time of screening, recorded in years. For participants aged 80 years or older, the age was recorded as 80.

### Distribution of Predictor Variables
The summary statistics presented in Figure 1 and the histograms of predictor variables shown in Figure 3 provide insights into the distribution of these variables:

- `Poverty`: The poverty index ranges from 0 to 5, with a mean of 3.077. The histogram indicates a marked left skew, suggesting that a significant proportion of participants fall into lower income categories.
- `Physical Activity Days (PhysActiveDays)`: The number of days participants engage in physical activity ranges from 2 to 7, with a mean value of 3.7 days per week.
- `Sleep Duration (SleepHrsNight)`:The number of hours participants sleep each night ranges from 2 to 12 hours, with a mean of 6.96 hours. The distribution of this variable approximates a normal curve.
- `Gender`: The gender distribution is nearly balanced, with 1,814 male participants and 1,759 female participants.
- `Age`: The age of participants spans from 16 to 80 years, with a mean age of 43.61 years. The histogram shows a slight right skew, with a concentration of participants aged between 29 and 56 years.

# Model

The goal of our modelling strategy is to use multilinear regression model to investigate the relationship between log(BMI) and poverty, Physical Activity Days, Sleep Duration, gender and age. 
Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $log(BMI)_i$ as the log(BMI). Then $\beta_i$ are the coefficients associated with each predictor variable, which represent the change in $log(BMI)_i$ for a one-unit change in the corresponding predictor, while holding all other predictors constant.

\begin{equation*} 
\log(\text{BMI}_i) = \beta_0 + \beta_1 \cdot \text{Poverty}_i + \beta_2 \cdot \text{PhysActiveDays}_i + \beta_3 \cdot \text{Age}_i + \beta_4 \cdot \text{SleepHrsNight}_i + \beta_5 \cdot \text{Gender}_i + \epsilon_i
\end{equation*}
\begin{equation*}
\epsilon_i \sim \text{Normal}(0, \sigma^2)
\end{equation*}

We ran the model in R [@citeR] using the lm() function for linear regression, with data manipulation performed using the dplyr package and data reading via the arrow package. No specific priors were applied, as this model relies on ordinary least squares (OLS) estimation, which assumes no prior distributions for the coefficients.

### Model justification
We use a multiple linear regression model to estimate the relationship between body mass index (BMI) and various predictors, including poverty index, physical activity days, age, sleep hours, and gender. The model is designed to predict log-transformed BMI, as this transformation helps normalize the distribution of BMI and address skewness. This approach allows us to explore how these lifestyle and socio-economic factors influence BMI across the sample, using high-quality survey data from the NHANES dataset. We employ predictors like poverty, physical activity, age, sleep hours, and gender to account for key factors that are known to influence BMI, based on existing research.

We use linear regression because it is appropriate for modeling the relationship between a continuous outcome variable (log(BMI)) and predictors physical activity, sleep hours, age, poverty, and gender. This method is straightforward to interpret, and the results could quantify the effect of predictors on BMI, making it a suitable choice for analyzing how these factors relate to BMI.

We chose to log-transform the BMI variable due to its right-skewed distribution. Log-transforming BMI allows us to better meet the assumptions of linear regression by stabilizing variance and making the relationship between BMI and predictors more linear. This is particularly important for ensuring that the model’s estimates are valid and interpretable.

We included the variable 'Poverty' to account for socio-economic status, as previous research has suggested that lower-income individuals tend to have higher BMI levels (Webber et al., 2023). 'PhysActiveDays' was included to capture the effect of physical activity on BMI, as increased physical activity is typically associated with a lower BMI(Webber et al., 2023). Age was included as a predictor to account for the natural changes in BMI that occur as individuals age. Sleep hours were added because there is evidence that insufficient sleep can contribute to weight gain (Ekstedt et al., 2013). Gender was included as a predictor due to well-established gender differences in BMI, with males typically having lower BMI than females (Longo-Silva et al., 2023).

We used the default settings in lm() from the stats package, which assumes normally distributed errors and does not apply any specific prior distributions, as this method does not require priors in the same way that Bayesian methods do. This approach provides a reliable way to assess the linear relationship between the predictors and BMI. However, we acknowledge that other methods, such as Bayesian regression or generalized linear models, could be used for more complex modeling, but this approach was selected for its simplicity and interpretability given the goals of the analysis.

Initially, we considered including interaction terms between predictors (such as between poverty and physical activity). However, when these interaction terms were added, the R-squared value decreased, indicating a reduced fit to the data. Additionally, the residual plot showed a noticeable concentration in the residuals, suggesting potential model misspecification or overfitting. As a result, we decided to exclude interaction terms to maintain a more interpretable and well-fitted model. Future analyses could revisit this approach with a larger dataset or alternative modeling strategies to better account for potential interactions.

By using this model, we aim to better understand the factors that influence BMI and provide insights into potential public health interventions that could target lifestyle changes and socio-economic factors to reduce BMI and related health risks.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-modelresults
#| tbl-cap: "Summary statistics for variables in the NHANES dataset."

# Load the saved linear model
lm_model <- readRDS("/Users/hxw/marriage/models/lm_model.rds")
# Tidy the model summary
model_summary <- tidy(lm_model)
# Extract R-squared value
r_squared <- summary(lm_model)$r.squared

# Add R-squared to the model summary table
model_summary_with_r2 <- model_summary %>%
  bind_rows(tibble::tibble(
    term = "R-squared",
    estimate = r_squared,
    std.error = NA,
    statistic = NA,
    p.value = NA
  ))

# Print the model summary table using knitr::kable
model_summary_with_r2 %>%
  kable(caption = "Summary of the Linear Model for Log(BMI)")
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-predictor-bmi
#| fig-cap: "Scatter plots of predictors vs. log(BMI) with fitted regression lines."
#| fig-subcap: ["Age vs. log(BMI)","Gender vs. log(BMI)", "Poverty vs. log(BMI)", "SleepHrsNight vs. log(BMI)","Physical Active Days vs. log(BMI)"]
#| layout-ncol: 3
#| fig.width: 8
#| fig.height: 12

# Scatter plot for Age vs. log(BMI)
plot_age_bmi <- ggplot(nhanes_data, aes(x = Age, y = log(BMI))) +
  geom_point(size = 0.3) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(x = "Age", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Box plot for Gender vs. log(BMI)
plot_gender_bmi <- ggplot(nhanes_data, aes(x = Gender, y = log(BMI))) +
  geom_boxplot(fill = "lightblue", color = "black") +  # Box plot
  labs(x = "Gender", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Scatter plot for Poverty vs. log(BMI)
plot_poverty_bmi <- ggplot(nhanes_data, aes(x = Poverty, y = log(BMI))) +
  geom_point(size = 0.3) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(x = "Poverty", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Scatter plot for SleepHrsNight vs. log(BMI)
plot_sleep_bmi <- ggplot(nhanes_data, aes(x = SleepHrsNight, y = log(BMI))) +
  geom_point(size = 0.3) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(x = "SleepHrsNight", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# Scatter plot for PhysActiveDays vs. log(BMI)
plot_physactive_bmi <- ggplot(nhanes_data, aes(x = PhysActiveDays, y = log(BMI))) +
  geom_point(size = 0.3) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  labs(x = "PhysActiveDays", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

plot_age_bmi
plot_gender_bmi
plot_poverty_bmi
plot_sleep_bmi
plot_physactive_bmi

```

# Discussion

## What is done in this paper? {#sec-first-point}

1.5 page This paper explores the relationship between Body Mass Index (BMI) and several key lifestyle and demographic factors, including poverty, physical activity, age, sleep duration, and gender. The analysis is performed using a linear regression model, which allows us to estimate how each of these predictors influences BMI, as measured by its log-transformed values. The dataset used for this study is the cleaned data from the US National Health and Nutrition Examination Survey (NHANES), which contains a broad range of health-related information from a representative sample of the U.S. population.

The model aims to quantify the impact of each predictor on BMI while controlling for the influence of other variables. The regression analysis reveals significant relationships between BMI and all the predictors considered, with some factors having a negative impact (e.g., poverty, physical activity) and others having a positive one (e.g., age, sleep hours, gender). This model sheds light on the ways in which various lifestyle choices and demographic factors may contribute to BMI variation across different segments of the population.

The model used in this study, a linear regression model, is a fundamental yet powerful tool that helps to identify the strength and direction of the relationships between predictors and BMI. The results provide a valuable starting point for understanding how changes in lifestyle and demographic factors may affect BMI. However, while the model is useful, it is not without its limitations. For instance, it assumes that the relationships between the predictors and BMI are linear, which may not always be the case. Additionally, the low R-squared value suggests that other important factors influencing BMI may not have been included in the model.

## What is something that we learn about the world?

1.5 page One of the key findings from this paper is that poverty is negatively correlated with BMI, and the relationship is statistically significant. The negative coefficient for the poverty variable suggests that individuals in poverty tend to have higher BMI, on average, than those with higher income levels. This result is consistent with a growing body of literature that links lower socioeconomic status to obesity and poor health outcomes. It may seem counterintuitive at first—since poverty is often associated with limited access to healthy foods and exercise—but the correlation can be explained by several factors.

In many low-income communities, the availability of affordable healthy food options is limited, leading individuals to rely on cheaper, calorie-dense foods that contribute to weight gain. Additionally, individuals in poverty may face multiple barriers to physical activity, including the lack of access to safe recreational spaces or the time and energy needed to engage in exercise, due to long working hours or other stressors. Furthermore, the high cost of healthcare and medical services in lower-income populations may prevent individuals from seeking preventative care and managing conditions like obesity. Thus, this finding underscores the importance of addressing socioeconomic disparities as part of efforts to combat obesity and improve public health.

Moreover, this finding highlights the need for policy interventions that promote access to healthy food and physical activity in low-income communities. It also suggests that any interventions aimed at reducing BMI must consider not just individual behaviors, but the broader socioeconomic context in which these behaviors occur. Policymakers should prioritize interventions that reduce the barriers to healthy living for individuals in poverty, such as improving access to nutritious food, increasing public health education, and providing community resources for physical activity.

## What is another thing that we learn about the world?

1.5 page Another key takeaway from this paper is that sleep duration has a significant negative correlation with BMI, suggesting that individuals who get more sleep tend to have lower BMI. This relationship is in line with previous research that has examined the link between sleep and obesity. The coefficient for sleep hours per night shows that for each additional hour of sleep, BMI decreases slightly, indicating that sleep plays a role in regulating body weight. This finding has important implications for public health, as sleep duration is a modifiable lifestyle factor that could be targeted in weight management programs.

The relationship between sleep and BMI can be explained through several mechanisms. Poor sleep has been shown to disrupt the body's metabolism, leading to an increase in appetite, particularly for high-calorie foods. Sleep deprivation also affects the hormones that regulate hunger and satiety, increasing the production of ghrelin (the hunger hormone) while decreasing the production of leptin (the satiety hormone). These changes can lead to overeating and, consequently, weight gain. Furthermore, insufficient sleep can lead to fatigue, reducing the likelihood of physical activity and further contributing to weight gain. Therefore, improving sleep hygiene and encouraging sufficient sleep may serve as an effective strategy for reducing BMI and addressing obesity.

This result also emphasizes the importance of taking a holistic approach to health. In addition to focusing on diet and physical activity, interventions aimed at improving sleep quality could be a key component in the prevention and treatment of obesity. Public health campaigns that educate individuals about the importance of sleep, along with strategies to improve sleep hygiene (e.g., creating a conducive sleep environment, avoiding caffeine, establishing regular sleep routines), could complement existing efforts to promote healthy eating and physical activity.

## Weaknesses and next steps

While the model results provide valuable insights into the relationship between log(BMI) and various predictor variables, there are some limitations that should be addressed in future analyses. One significant weakness is the relatively low R-squared value of 0.02335, indicating that the model only explains a small portion of the variance in BMI. This suggests that there are other factors, potentially unaccounted for in the model, that influence BMI. Although several predictors show statistically significant relationships with BMI, the explanatory power of the model remains limited. It is possible that additional variables, such as dietary habits, genetics, or environmental factors, could improve the model's ability to explain the variation in BMI.

Another limitation is the potential for omitted variable bias. Despite including a range of relevant predictors, there may be other important factors that have not been incorporated into the model. For instance, socioeconomic status, mental health, or access to healthcare may play significant roles in determining BMI but were not included in the current analysis. Future models could benefit from a more comprehensive selection of predictor variables to provide a more holistic understanding of BMI variation.

Additionally, the model assumes linear relationships between the predictors and BMI, which may not fully capture the complexities of the data. Interaction effects, such as between poverty and physical activity, could potentially reveal more nuanced relationships, but these were not included in the current model. The absence of interaction terms may have led to the underestimation of the impact of certain predictors, as the effect of one variable may depend on the level of another. Future analyses could explore interaction terms to assess whether these improve model fit and offer deeper insights into the factors that influence BMI.

Moving forward, there are several steps that could be taken to address these weaknesses. First, the inclusion of additional predictors could improve the model's explanatory power. Collecting more detailed data on lifestyle, diet, or mental health factors would help provide a more complete picture of the determinants of BMI. Second, using a non-linear model or exploring transformations of the predictors might better capture the relationships between BMI and the explanatory variables. Finally, addressing the possibility of interaction effects and testing different model specifications could enhance the robustness of the results. This could include experimenting with non-parametric models, such as random forests or gradient boosting machines, to better handle the complexities of the data. By expanding the scope of the analysis and refining the model, future work could yield more precise estimates and offer more actionable insights for public health interventions.

\newpage

\appendix

# Appendix {-}

# Additional data details {#sec-data-details}

## Data Justication
The NHANES_raw dataset provides a comprehensive collection of data from the American National Health and Nutrition Examination Surveys (NHANES), including 20,293 observations spanning 75 variables. This dataset includes additional variables that describe the sample weighting scheme employed to account for the complex survey design. However, despite its breadth and richness, NHANES_raw introduces complexities in analysis due to the oversampling of certain subpopulations, such as racial minorities. This sampling approach necessitates the use of survey weights and other design-specific parameters to obtain unbiased estimates and valid statistical inferences.

Naïve analysis of NHANES_raw—without accounting for the sample design and weights—can lead to significant biases and erroneous conclusions. For instance, the racial composition in NHANES_raw does not represent the true demographic distribution of the U.S. population due to oversampling adjustments. As a result, researchers must account for these weights in any analysis to avoid overstating significance levels and generating inaccurate results. This makes NHANES_raw unsuitable for straightforward analysis in contexts where the primary goal is to demonstrate simple statistical relationships, especially for educational purposes.

In contrast, the NHANES dataset, which is a resampled subset of NHANES_raw, addresses the complexities associated with the oversampling design. With 10,000 observations and the same 75 variables, NHANES is simplified to approximate a simple random sample from the U.S. population. This resampling process removes the need for researchers to manually account for the oversampling effects, enabling analyses that are easier to interpret and less prone to methodological errors for educational purposes.

The decision to use NHANES in this paper was driven by its appropriateness for demonstrating relationships between body mass index (BMI) and lifestyle, poverty, age, and gender. NHANES allows for clear and reliable analyses without requiring specialized expertise in survey design or sampling weights, making it ideal for this study’s educational focus. The dataset enables meaningful exploration of public health questions while minimizing the risk of misinterpretation due to unaddressed biases in the sampling framework.

It is important to acknowledge that NHANES is adapted for educational purposes and is not intended for formal research. Researchers conducting advanced studies should rely on the original NHANES_raw data and adhere to the prescribed sampling and weighting procedures to ensure the validity of their findings. Future research could extend the current analysis by using NHANES_raw in combination with survey weighting techniques to explore population-level health trends and disparities with greater accuracy.

# Surveys, Sampling, and Observational Data {#sec-sampling-details}

# Model details {#sec-model-details}

## Diagnostics

The analysis of linear regression assumptions shows that all conditions are met. The fitted versus residual plot (Figure 8) displays a null plot centered around zero, indicating that the linearity assumption is satisfied, as no discernible pattern exists. Additionally, the consistent spread of residuals across the fitted values confirms the constant error variance assumption. The absence of correlation or patterns suggests that the independence of errors is also satisfied. Both the fitted versus standardized residuals plot (Figure 9) and the predictor variables versus residuals plots (Figures 11 to 13) support these conclusions.
The QQ plot (Figure 15) shows standardized residuals following a straight diagonal line, while the histogram (Figure 10) indicates a distribution close to ~N(0,1), confirming the normality assumption of errors. Lastly, the response variable versus fitted values plot (Figure 14) shows that observed values align with predicted values, suggesting that the model accurately predicts the response variable.
Although the residual plots indicate that all regression assumptions are met, the predictor versus response plots (Figures 16 to 20) show weak correlations between predictor variables and log(BMI), with log(BMI) approximated by

```{r}
#| echo: false
#| label: fig-model-diagnostics
#| fig-cap: "Diagnostic and Predictor Relationships in the Linear Model for log(BMI)."
#| fig-subcap: ["Fitted vs Residuals", "Fitted vs Standardized Residuals", "Histogram of Standardized Residuals", "Poverty vs Residuals", "Age vs Residuals", "Gender vs Residuals", "Fitted vs log(BMI)", "Normal Q-Q Plot"]
#| layout-ncol: 4
#| fig.width: 12
#| fig.height: 18

# Extract values for residuals and fitted values
fitted_values <- fitted(lm_model)
residual_values <- resid(lm_model)
sresidual_values <- rstandard(lm_model)

# Data for scatter plots
scatter_data <- nhanes_data %>%
  mutate(log_BMI = log(BMI))

# Create each plot
# 1. Fitted vs Residuals
plot_fitted_residuals <- ggplot(data.frame(fitted_values, residual_values), 
                                aes(x = fitted_values, y = residual_values)) +
  geom_point(color = "blue", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Fitted Values", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 2. Fitted vs Standardized Residuals
plot_fitted_standardized <- ggplot(data.frame(fitted_values, sresidual_values), 
                                   aes(x = fitted_values, y = sresidual_values)) +
  geom_point(color = "red", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 3. Standardized Residuals Histogram
plot_residuals_hist <- ggplot(data.frame(sresidual_values), aes(x = sresidual_values)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(x = "Standardized Residuals", y = "Frequency") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 4. Residuals vs Poverty
plot_residuals_poverty <- ggplot(data.frame(nhanes_data$Poverty, residual_values), 
                                 aes(x = nhanes_data$Poverty, y = residual_values)) +
  geom_point(color = "green", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Poverty Index", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 5. Residuals vs Age
plot_residuals_age <- ggplot(data.frame(nhanes_data$Age, residual_values), 
                             aes(x = nhanes_data$Age, y = residual_values)) +
  geom_point(color = "orange", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Age", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 6. Residuals vs Gender
plot_residuals_gender <- ggplot(data.frame(nhanes_data$Gender, residual_values), 
                                aes(x = nhanes_data$Gender, y = residual_values)) +
  geom_boxplot(fill = "gray") +
  labs(x = "Gender", y = "Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 7. Fitted vs Log(BMI)
plot_fitted_logBMI <- ggplot(data.frame(fitted_values, log_BMI = scatter_data$log_BMI), 
                             aes(x = fitted_values, y = log_BMI)) +
  geom_point(color = "purple", size = 0.3) +
  geom_abline(intercept = 0, slope = 1, color = "blue", linetype = "dashed") +
  labs(x = "Fitted Values", y = "log(BMI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

# 8. Q-Q Plot
plot_qq <- ggplot(data = data.frame(std_resid = rstandard(lm_model)), aes(sample = std_resid)) +
  stat_qq(color = "darkred", size = 0.3) +
  stat_qq_line(color = "blue") +
  labs(x = "Theoretical Quantiles", y = "Standardized Residuals") +
  theme_minimal() +
  theme(axis.title = element_text(size = 10), axis.text = element_text(size = 8))

plot_fitted_residuals
plot_fitted_standardized
plot_residuals_hist
plot_residuals_poverty
plot_residuals_age
plot_residuals_gender
plot_fitted_logBMI
plot_qq

```

\newpage

# References